The Astropy Project: \\ Sustaining and Growing a Community-oriented Open-source Project and the Latest Major Release (v5.0) of the Core Package

### 2022-01-26 src/ms.tex
Fix subpackage names
astropy roadmap

%Briefly summarize new procedures and governance structure, new CoCo, election
%process overview, etc.

As part of the process of developing Astropy into a long-term sustainable
product, and to improve transparency and accountability, the Project agreed to
write down and formalize our governance structure (partly supported by explicit
funding for this purpose - see \S \ref{sec:project-funding}). At the 2019
coordination meeting input was gathered from participants on what governance
structures existed in the associated Open Source Software communities, and what
would fit well with the needs of Astropy. This led into a ``retreat'' planned
for March 2020, but due to the COVID-19 pandemic, this became a serious of
virtual meetings of the ``Astropy Governance Working Group''.  This group
drafted the APE0 document \citep{ape0}, which was then eventually ratified and
implemented by the ``Astropy Governance Implementation Working Group'' in Fall
2021. While the process emphasized flexibility and the ability to adapt to
changing circumstances, it is expected that this is the framework Astropy's
governance will operate in for at least the medium-term future.

The APE0 \citep{ape0} document lays out the principles of this governance
structure, so we refer the reader to that document for a more thorough
description.  However here we highlight some key elements. While many of these
principles were already de facto true or have been discussed organically (and
have been discussed in earlier papers in this series), the APE0-based governance
aimed to provide a single place where the community can agree as a starting
point. With this in mind, it highlights the developer and user community of the
Astropy Project as the ultimate sources of authority, as well as the core
principle of ``do-ocracy'' that those who do work for the Project (be it coding,
training, or other less concrete contributions) gain more influence on the
outputs of the Project by virtue of their effort.  APE0 adds, however, the
concept of ``voting members'' - a self-governed part of that community who are
entrusted to elect the Coordination Committee. While this committee has existed
from the inception of the project, APE0 establishes a formal voting process for
this committee, and explicitly outlines the rights and responsibilities of the
Coordination Committee. This role is mainly to facilitate consensus and act as
the decision maker when other mechanisms have failed.  However, it also includes
powers that either require central authority or secrets (e.g. passwords), but
AOE0 also charges the Committee to devolve responsibilities and seek community
input on these items as often as possible.

The first coordinating committee election under these rules took place in fall
2021, electing a mix of prior existing and new coordination committee members,
and was contested in the sense of more candidates than available slots. This
suggests the process is already working to serve the long-term interests of the
committee to both spread the coordination effort, and to ensure it is not
dominated by the same people for as long as the Project continues. While other,
more fine-grained governance improvements are planned for the future, it is
clear the foundation is now in place.

@@

Overview of roadmap and some highlights.

>>>>

Briefly summarize new procedures and governance structure, new CoCo, election
process overview, etc.

@@

The Astropy Roadmap is a document in the astropy-project repository that captures high level actionable items that the Astropy project aims to undertake to improve the health and stability of the  project. It is a static document that is revisited regularly at the Astropy coordination meetings, to keep track of progress and write new versions as needed. There is a related project board linked to the Roadmap document, that holds specific issues and efforts related to Roadmap items. The project board is a living document that is continually updated as work is planned, assigned, and completed.

All items in the Roadmap have been agreed to be priorities for the Astropy Project, and are color-coded based on resources (both time/effort and developers/expertise) needed to complete the item. Green items are well underway, have sufficient resources/support and a plan in place for completion. Orange items are well defined, and work for aquiring sufficient resources underway. Red items do not yet have a plan for implementation and need more resources.

The Astropy Roadmap originates from the March 2021 Astropy Coordination meeting where it was first drafted before being handed off to a newly formed Astropy Roadmap working group for completion. The Roadmap in its currant form was adopted via pull request in December 2021.
